Retrofitting Effect Handlers onto OCaml

### 2020-11-12 retro-concurrency.tex
abstract

\begin{abstract}
  Effect handlers have gained immense interest in the recent years as a means
for modular programming with user-defined effects. The runtime semantics of
effect handlers is particularly attractive as they allow advanced control flow
mechanisms such as first-class continuations, generators, async/await,
coroutines to be composably expressed. This paper presents a design, a
full-fledged implementation and an evaluation of effect handlers for OCaml.
OCaml is industrial-strength multi-paradigm programming language that tends to
be widely used for systems programming, our implementation strives to maintain
the existing performance profile and toolchain compatibility. Retrofitting
effect handlers onto OCaml is challenging since OCaml does not have any
non-local control flow mechanisms other than exceptions. Our implementation of
effect handlers for OCaml imposes negligible overhead on code that does not use
effect handlers, remains compatible with program analysis tools that inspect the
stack and is efficient for code that makes use of effect handlers.
\end{abstract}

>>>>

\begin{abstract}
Effect handlers have been gathering momentum as a mechanism
for modular programming with user-defined effects. The runtime semantics of
effect handlers allow for control flow
mechanisms such as first-class continuations, generators, async/await and coroutines to be composably expressed. We present a design and evaluate a
full-fledged efficient implementation of effect handlers for OCaml, an industrial-strength multi-paradigm programming language. Our implementation strives to maintain the existing performance profile of conventional OCaml code, particularly in the domain of systems and toolchain programming. Retrofitting effect handlers onto OCaml is challenging since OCaml does currently not have any non-local control flow mechanisms other than exceptions. Our implementation of
effect handlers for OCaml: {\em (i)} imposes negligible overhead on code that does not use
effect handlers; {\em (ii)} remains compatible with program analysis tools that inspect the
stack; and {\em (iii)} is efficient for new code that makes use of effect handlers.
\end{abstract}
